00-Data-Processing-Example

### 2020-11-13 examples/00-Data-Processing-Example.Rmd
v1 Updated 00-Data-Processing-Example.Rmd

@@

@@

Plot the DAs and the service areas:
```{r}
ggplot() +
  geom_sf(data = hamilton_da_2016,
          color = "lightgray") +
  geom_sf(data = sobi_service,
          fill = NA)
```

We want to work only with the DAs that are within a reasonable buffer of the service areas (e.g., 500 m):
```{r}
ggplot(hamilton_da_2016[sobi_service %>% st_buffer(500),]) +
  geom_sf(color = "lightgray") + 
  geom_sf(data = sobi_service,
          fill = NA)
```

Extract DAs in service areas:
```{r}
hamilton_da_sobi <- hamilton_da_2016[sobi_service %>% st_buffer(500),]
```

Extract parks in service areas:
```{r}
parks_sobi <- parks[sobi_service %>% st_buffer(500),]
```

@@

Plot population:
```{r}
ggplot() +
  geom_sf(data = hamilton_da_sobi,
          aes(fill = population),
          color = NA) +
  geom_sf(data = sobi_service,
          fill = NA) +
  geom_sf(data = parks_sobi,
          fill = "green") +
  scale_fill_distiller(palette = "YlOrRd", direction = 1)
```

Remove the parks from the DAs (use `st_difference()`):
```{r}
# First combine `parks_sobi` to generate a single geography object, and make valid
parks_sobi <- st_combine(parks_sobi) %>% 
  st_make_valid()
# The new simple features object will be the difference of the two objects
hamilton_da_sobi_clean <- st_difference(hamilton_da_sobi, parks_sobi)
```

Plot the new DAs (minus parks) and compare to original DAs:
```{r}
ggplot() +
  geom_sf(data = hamilton_da_sobi_clean,
          color = NA,
          fill = "blue") +
  geom_sf(data = hamilton_da_sobi,
          fill = NA,
          color = "black")
```

@@

@@

Extract population cells in the service area (and buffer)
```{r}
interpolated_pop <- interpolated_pop[sobi_service %>% st_buffer(500),]
```

Plot interpolated population again:
```{r}
ggplot() + 
  geom_sf(data = interpolated_pop,
          aes(color = population)) +
  geom_sf(data = hamilton_da_sobi,
          fill = NA,
          color = "lightgray") + 
  geom_sf(data = sobi_service,
          fill = NA) +
  coord_sf(xlim = )
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1)
```

Bounding box of population in serviced areas:
```{r}
bounding_box <- st_bbox(interpolated_pop)
```

Plot:
```{r}
ggplot() + 
  geom_sf(data = interpolated_pop,
          aes(color = population)) +
  geom_sf(data = hamilton_da_sobi,
          fill = NA,
          color = "lightgray") + 
  geom_sf(data = sobi_service,
          fill = NA) +
  geom_sf(data = st_as_sfc(bounding_box),
          linetype = "dashed",
          fill = NA) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1)
```

Select population within bounding box:
```{r}
interpolated_pop <- interpolated_pop[st_as_sfc(bounding_box),]
```

Plot:
```{r}
ggplot() + 
  geom_sf(data = interpolated_pop,
          aes(color = population)) +
  geom_sf(data = hamilton_da_sobi,
          fill = NA,
          color = "lightgray") + 
  geom_sf(data = parks,
          color = NA,
          fill = "lightgreen") +
  geom_sf(data = sobi_service,
          fill = NA) +
  geom_sf(data = st_as_sfc(bounding_box),
          linetype = "dashed",
          fill = NA) +
  coord_sf(xlim = c(bounding_box[1], bounding_box[3]), # min & max of x values
           ylim = c(bounding_box[2], bounding_box[4])) + # min & max of y values
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1)
```

>>>>

Bounding box of population in serviced areas:
```{r}
bounding_box <- st_bbox(sobi_service %>%
                          st_buffer(500))
```

Plot:
```{r}
ggplot() + 
  geom_sf(data = hamilton_da_2016,
          fill = NA,
          color = "lightgray") + 
  geom_sf(data = sobi_service,
          fill = NA) +
  geom_sf(data = st_as_sfc(bounding_box),
          linetype = "dashed",
          fill = NA)
```

Plot:
```{r}
ggplot() + 
  geom_sf(data = hamilton_da_2016,
          fill = NA,
          color = "lightgray") + 
  geom_sf(data = sobi_service,
          fill = NA) +
  geom_sf(data = st_as_sfc(bounding_box),
          linetype = "dashed",
          fill = NA)
```

@@

Read cemeteries:
```{r}
cemeteries <- st_read("Cemeteries.shp")
```

Read environmentally sensitive areas:
```{r}
esa <- st_read("Environmentally_Sensitive_Areas_Boundaries.shp")
```

@@

Extract DAs in service areas:
```{r}
hamilton_da_sobi <- hamilton_da_2016[st_as_sfc(bounding_box),]
```

Extract parks in service areas:
```{r}
parks_sobi <- st_crop(parks, st_as_sfc(bounding_box))
```

Extract cemeteries in service areas:
```{r}
cemeteries_sobi <- st_crop(cemeteries, st_as_sfc(bounding_box))
```

Extract environmentally sensitive areas in service areas:
```{r}
esa_sobi <- st_crop(esa, st_as_sfc(bounding_box))
```

@@

Plot population:
```{r}
ggplot() +
  geom_sf(data = hamilton_da_sobi,
          aes(fill = population),
          color = NA) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_fill_distiller(palette = "YlOrRd", direction = 1)
```

Plot "Green" spaces:
```{r}
ggplot() +
  geom_sf(data = parks_sobi,
          fill = "blue") +
  geom_sf(data = cemeteries_sobi,
          fill = "gray") +
  geom_sf(data = esa_sobi,
          fill = "green") +
  geom_sf(data = sobi_service,
          fill = NA,
          color = "black") +
  geom_sf(data = hamilton_da_sobi,
          fill = NA,
          color = "gray") +
  scale_fill_distiller(palette = "YlOrRd", direction = 1)
```

Union of the "green space" features:
```{r}
parks_sobi <- st_combine(parks_sobi)
cemeteries_sobi <- st_combine(cemeteries_sobi)
esa_sobi <- st_combine(esa_sobi)
```

Check that the topology is valid:
```{r}
st_is_valid(parks_sobi)
st_is_valid(cemeteries_sobi)
st_is_valid(esa_sobi)
```

Make valid:
```{r}
parks_sobi <- st_make_valid(parks_sobi)
esa_sobi <- st_make_valid(esa_sobi)
```

Remove the parks from the DAs (use `st_difference()`):
```{r}
hamilton_da_sobi_clean <- st_difference(hamilton_da_sobi, parks_sobi)
```

Remove the cemeteries from the DAs (use `st_difference()`):
```{r}
hamilton_da_sobi_clean <- st_difference(hamilton_da_sobi_clean, cemeteries_sobi)
```

Remove environmentally sensitive areas from the DAs (use `st_difference()`):
```{r}
hamilton_da_sobi_clean <- st_difference(hamilton_da_sobi, esa_sobi)
```

Plot the new DAs (minus green) and compare to original DAs:
```{r}
ggplot() +
  geom_sf(data = hamilton_da_sobi_clean,
          color = NA,
          fill = "blue") +
  geom_sf(data = hamilton_da_sobi,
          fill = NA,
          color = "black")
```

@@

Check the population totals:
```{r}
sum(hamilton_da_sobi_clean$population)
sum(interpolated_pop$population)
```

@@

Extract population cells in the service area (and buffer)
```{r}
interpolated_pop_clean <- interpolated_pop[sobi_service %>% st_buffer(500),]
```

Plot interpolated population again:
```{r}
ggplot() + 
  geom_sf(data = interpolated_pop_clean,
          aes(color = population)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1)
```
